---
title: "ISLR 3: Linear Regression"
author: "Justin M Shea"
date: ''
output:
  pdf_document:
        toc: TRUE
  html_document: default
---

\newpage



Load the following packages.
```{r}
library(MASS)
library(ISLR)
```

## Simple linear regression
Check out the `Boston` data
```
?Boston
```
```{r}
names(Boston)
class(Boston)
dim(Boston)
```
```{r}
summary(Boston)
```
Lets plot the `Boston` data
```{r}
plot(medv ~ lstat, data = Boston)
```

Run a linear model (`lm`) on it and print the results

```{r}
Boston_lm <- lm(medv ~ lstat, data = Boston)
Boston_lm
```

Lets plot the linear model against a scatter plot of `medv` and `lstat`.

```{r}
plot(medv ~ lstat, data = Boston)
abline(Boston_lm, col="red")
```

We can check the confidence intervals of our models parameters using the `confint` function.
```{r}
confint(Boston_lm)
```

In addition, the `predict` function is useful in making some predictions with the `Bonston_lm` model we created.
```{r}
predict(Boston_lm, data.frame(lstat=c(5,10,15)), interval="confidence")
```

\newpage

## Multiple linear regression

Regress `lstat` and `age` against `medv` of the `Boston` data set and print the `summary` diagnostics. 

```{r}
Boston_lm2 <- lm(medv ~ lstat + age, data = Boston)
summary(Boston_lm2)
```

Use the `.` notation to regress all variables in the `Boston` data against `medv` and print the `summary` diagnostics.
```{r}
Boston_lm3 <- lm(medv ~., data = Boston)
summary(Boston_lm3)
```

Lets plot the model
```{r}
par(mfrow=c(2,2))
plot(Boston_lm3)
```

Use the `.` notation again to select all variables, but this time subtract `age` and `indus` before regressing against `medv`. Print the `summary` diagnostics.

```{r}
Boston_lm4 <- update(Boston_lm3,~.-age-indus)
summary(Boston_lm4)
```

## Nonlinear terms and Interactions

Multiply `lstat` by `age` and regress against `medv`:

```{r}
Boston_mult <- lm(medv ~ lstat * age, data = Boston)
Boston_mult
```

Create a quadratic interaction using the `I` function with `lstate`:
```{r}
Boston_Interaction <- lm(medv ~ lstat + I(lstat^2), data = Boston)
```

Create a $4^{th}$ order polynomial wrapping `lstat` in the `poly` function, defining the degree arguments a `4`.
```{r}
Boston_poly <- lm(medv ~ poly(lstat, degree=4), data = Boston)
```

Plot the results of the both the `Boston_Interaction` and `Boston_poly` models.
```{r}
attach(Boston)
par(mfrow=c(1,1))
plot(medv ~ lstat, data = Boston)
points(lstat, fitted(Boston_Interaction), col="red", pch=20)
points(lstat, fitted(Boston_poly), col="blue", pch=20)
```

Finally, lets look at the coeffients of both:
```{r}
Boston_Interaction
Boston_poly
```

## Qualitative predictors

For this section, use the `Carseats` data. Lets explore:

```{r, eval = FALSE}
?Carseats
```
```{r}
names(Carseats)
summary(Carseats)
```

Run a model regressing all variables as well as two new interactive variables resulting form combining `Income:Adversting` and `Age:Price`.
```{r}
Carseats_lm <- lm(Sales ~. + Income:Advertising + Age:Price, data = Carseats)
summary(Carseats_lm)
```

Call the `contrasts` function on the `ShelveLoc` variable to display a table of levels corresponding to the quality of the shelving location for the car seats at each site.
```{r}
contrasts(Carseats$ShelveLoc)
```


## Brief section writing R functions

Function creating a plot which displays linear model regressin line.

```{r}
regplot <- function(x, y){
          fit <- lm(y ~ x)
          plot(x, y)
          abline(fit, col="red")
}

attach(Carseats)
regplot(Price, Sales)
```

This time, add the `...` argument to the function, which allowing one to pass arguements to functions within the function.
```{r}
regplot <- function(x, y, ...){
          fit <- lm(y ~ x)
          plot(x, y, ...)
          abline(fit, col="red")
}

regplot(Price, Sales, xlab="Price", ylab="Sales", col="blue", pch=20)

```