---
title: "Model Selection"
author: "Justin M Shea"
date: ''
output:
  pdf_document: default
  html_document: default
---

Load the `ISLR` package and check the the `Hitters` data.

```{r}
library(ISLR)
data(Hitters)
```
```{r, eval=FALSE}
?Hitters
```
```{r}
str(Hitters)
```

Are there any missing values?
```{r}
NA_index <- is.na(Hitters)
length(Hitters[NA_index])
```

There are `r length(Hitters[NA_index]` missing values here, so before we proceed we will remove them:

```{r}
Hitters <- na.omit(Hitters)

NA_index <- is.na(Hitters)

length(Hitters[NA_index])
```

## Best Subset regression

We will now use the package `leaps` to evaluate all the best-subset models.

```{r}
library(leaps)
subset_full <- regsubsets(Salary ~ ., data = Hitters)
summary(subset_full)
```

Notice above, the default best-subsets up to size 8.
Lets increase that to 19, which is all the variables, create `summary` statistics on the model and view their names.

```{r}
subset_full <- regsubsets(Salary ~ ., data = Hitters, nvmax = 19)

full_summary <- summary(subset_full)

names(full_summary)
```

So lets plot the `Cp`, or the estimated prediction error, for each variable.

```{R}
plot(full_summary$cp, xlab = "Number of Variables", ylab = "Cp")
which.min(full_summary$cp)
points(10, full_summary$cp[10], pch = 20, col = "red")
```

There is a plot method designed specifically for the `regsubsets`  object which is displayed below. It also plots the `Cp` statistic but each variable. Areas that are colored black indicate the variable is present in the model at the corresponding `Cp` level, while white areas communicate an absence of the variable.

```{r}
plot(subset_full, scale = "Cp")
coef(subset_full, 10)
```



## Forward Stepwise Selection

Here we use the `regsubsets` function but specify the `method="forward"` option:

```{r}
forward_step <- regsubsets(Salary ~ ., data=Hitters, nvmax=19, method="forward")
summary(forward_step)
plot(forward_step, scale = "Cp")
```

## Model Selection Using a Validation Set

Lets make a training and validation set, so that we can choose a good subset model. We will do it using a slightly different approach from what was done in the the book.

```{r}
dim(Hitters)
set.seed(1)
train <- sample(seq(263), 180, replace=FALSE)
train
forward_step <- regsubsets(Salary ~ ., data=Hitters[train,], nvmax=19, method="forward")
```

Now we will make predictions on the observations not used for training. We know there are 19 models, so we set up some vectors to record the errors. We have to do a bit of work here, because there is no predict method for `regsubsets`.

```{r}
val.errors <- rep(NA, 19)
x.test <- model.matrix(Salary ~ ., data = Hitters[-train,]) # notice the -index!

for(i in 1:19){
  coefi <- coef(forward_step, id = i)
  pred <- x.test[,names(coefi)] %*% coefi
  val.errors[i] <- mean((Hitters$Salary[-train] - pred)^2)
}

plot(sqrt(val.errors), ylab="Root MSE", ylim=c(300,400), pch=19, type="b")
points(sqrt(forward_step$rss[-1]/180), col="blue", pch=19, type="b")
legend("topright", legend=c("Training", "Validation"),col=c("blue", "black"), pch=19)
```

As we expect, the training error goes down monotonically as the model gets bigger, but not so for the validation error.

This was a little tedious - not having a predict method for `regsubsets`. So we will write one!
```{r}
predict.regsubsets <- function(object,newdata, id, ...){
                        form <- as.formula(object$call[[2]])
                        mat <- model.matrix(form, newdata)
                        coefi <- coef(object, id = id)
                        mat[,names(coefi)] %*% coefi
}
```


## Model Selection by Cross-Validation
 

We will do 10-fold cross-validation. Its really easy!

```{r}
set.seed(11)
folds <- sample(rep(1:10, length=nrow(Hitters)))
folds
table(folds)

cv.errors <- matrix(NA, 10, 19)

for(k in 1:10){
best.fit <- regsubsets(Salary ~ ., data = Hitters[folds!=k,], nvmax=19, method="forward")
  for(i in 1:19){
                pred <- predict(best.fit, Hitters[folds==k,], id = i)
                cv.errors[k, i] <- mean((Hitters$Salary[folds==k] - pred)^2)
  }
}

rmse.cv <- sqrt(apply(cv.errors, 2, mean))

plot(rmse.cv, pch=19, type="b")
```


## Ridge Regression and the Lasso
 
We will use the package `glmnet`, which does not use the model formula language, so we will set up an `x` and `y`.

```{r}
library(glmnet)
x <- model.matrix(Salary ~ .-1, data = Hitters) 
y <- Hitters$Salary
```

First we will fit a ridge-regression model. This is achieved by calling `glmnet` with `alpha=0` (see the help file). There is also a `cv.glmnet` function which will do the cross-validation for us. 

```{r}
fit.ridge <- glmnet(x, y, alpha = 0)
plot(fit.ridge, xvar= "lambda", label = TRUE)
cv.ridge <- cv.glmnet(x, y, alpha = 0)
plot(cv.ridge)
```

Now we fit a lasso model; for this we use the default `alpha=1`

```{r}
fit.lasso <- glmnet(x,y)
plot(fit.lasso,xvar="lambda",label=TRUE)

cv.lasso <- cv.glmnet(x,y)
plot(cv.lasso)
coef(cv.lasso)
```

 Suppose we want to use our earlier train/validation division to select the `lambda` for the lasso.This is easy to do.
 
```{r}
lasso.tr <- glmnet(x[train,],y[train])
lasso.tr

pred <- predict(lasso.tr,x[-train,])
dim(pred)
rmse <-  sqrt(apply((y[-train]-pred)^2,2,mean))
plot(log(lasso.tr$lambda),rmse,type="b",xlab="Log(lambda)")

lam.best <- lasso.tr$lambda[order(rmse)[1]]
lam.best
coef(lasso.tr,s=lam.best)
```
